---
title: "Assignment_3"
author: "Meenakshi Vaidhiyanathan"
date: "2024-03-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing required packages including `naivebayes` package
```{r}
#install.packages("reshape")
#install.packages("reshape2")
#install.packages("melt")
#install.packages("naivebayes")
#install.packages("pROC")
```

loading the necessary libraries 
```{r}
library(readr)
library(dplyr)
library(caret)
library(class)
library(melt)
library(reshape)
library(reshape2)
library(ggplot2)
library(ISLR)
library(naivebayes)
library(e1071)
library(pROC)
```


Importing the dataset and assigning it to the variable `bank`. Using the head function to display first six rows of the dataset `bank`.
```{r}
bank <- read_csv("~/Downloads/UniversalBank.csv")
head(bank)
```

The is.na() function is used to check if there are any missing values in the dataset and is assigned to `bank_na` variable. 
```{r}
bank_na <- is.na.data.frame("universalbank")
```


The column Online, Credit card and personal loan have categorical variables which are converted to factors using the `as.factor()` function. 
```{r}
bank$Online<- as.factor(bank$Online)
bank$CreditCard<- as.factor(bank$CreditCard)
bank$`Personal Loan`<- as.factor(bank$`Personal Loan`)
```

#Data Partition and Normalization
The process of Data partition is performed to create the training data that is asssigned to `Train_data` containing 60% of the dataset `bank` and the Validation data assigned to `Valid_data` containing the remaining 40% of the dataset `bank`.  
```{r}
set.seed(333)
Train_Index<- createDataPartition(bank$`Personal Loan`, p=0.6, list=FALSE)
Train_data <-bank[Train_Index,]
Valid_data <-bank[-Train_Index,]
```


Both Training and validation data is normalised using the fucntions `preProcess()` and `predict()` and assinged to `Train_normalised` and `Valid_normalised` respectively. 
```{r}
Model_normalised <- preProcess(Train_data[,-c(10,13:14)],method = c("center", "scale"))
Train_normalised <- predict(Model_normalised,Train_data)
Valid_normalised<- predict(Model_normalised,Valid_data)
```


The following code chink creates a pivot table for training data using the `table()` function. 
```{r}
Table.OCP <- table(Train_normalised$`Personal Loan`, Train_normalised$Online, Train_normalised$CreditCard, dnn=c("Personal Loan", "Online", "Credit Card"))
Table.OCP
```


Part B: Computing P(Loan | Online & CC)
As we look that the pivot table created in part A out of the total 550 records where of active online banking users with credit cards, 49 had accepted a personal loan, so $$\textbf{P}(\text{Loan}=1\ |\ \text{CC}=1\text {and}\text{ Online}=1) = \frac{49}{550} = 0.089$$.


Computing P(loan | Online & CC)
```{r}
Table.OCP[2,2,2] / (Table.OCP[2,2,2] + Table.OCP[1,2,2])
```


Part C: Creating two separate pivot tables for Training data.One will have Loan as rows which is as a function of Online i.e columns and the other will have Loan (rows) as a function of CC. The `table() function is used.
```{r}
online_table <- table(Train_normalised$`Personal Loan`, Train_normalised$Online, dnn=c("Personal Loan", "Online"))
online_table

credit_table <- table(Train_normalised$`Personal Loan`, Train_data$CreditCard, dnn=c("Personal Loan", "Credit Card"))
credit_table
```


Part D : Computing the following quantities: 
i)P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)
i) $$\textbf{P}\text{(CC = 1 | Loan = 1)}\text{= 80/80+208}$$

```{r}
probability_ccl <- credit_table[2,2] / (credit_table[2,2] + credit_table[2,1])
probability_ccl
```
ii) P(Online = 1 | Loan = 1)
ii)$$\textbf{P}\text{(Online = 1 | Loan = 1)}\text{= 184/184+104}$$

```{r}
probability_ol <- online_table[2,2] / (online_table[2,2] + online_table[2,1])
probability_ol
```


iii)P(Loan = 1) (the proportion of loan acceptors)
iii) $$\textbf{P}\text{(Loan = 1)}\text{= 288/288+2712}$$

```{r}
probability_loan <- sum(Train_normalised$`Personal Loan`==1) / length(Train_normalised$`Personal Loan`)
probability_loan
```

iv) P(CC = 1 | Loan = 0)
iv) $$\textbf{P}\text{(CC = 1 | Loan = 0)}\text{= 806/806+1906}$$

```{r}
probability_ccnl <-credit_table[1,2] / (credit_table[1,2] + credit_table[1,1])
probability_ccnl
```

v)P(Online = 1 | Loan = 0)
v) $$\textbf{P}\text{(Online = 1 | Loan = 0)}\text{= 1621/1621+1091}$$

```{r}
probability_onl <- online_table[1,2] / (online_table[1,2] + online_table[1,1])
probability_onl
```

vi)P(Loan = 0)
vi) $$\textbf{P}\text{(Loan = 0)}\text{= 2712/2712+288}$$

```{r}
probability_nl <- sum(Train_normalised$`Personal Loan`==0) / length(Train_normalised$`Personal Loan`)
probability_nl
```

Part E : The computed quantities from above were used for the Naive Bayes probability P(Loan = 1 | CC = 1, Online = 1).
$$\textbf{P}(\text{Loan}=1\ |\ \text{CC}=1,\text{ Online}=1)\text{= (0.6388 x 0.2777 x 0.096) / (0.6388 x 0.2777 x 0.096 + 0.5977 x 0.2972 x 0.904)= 0.0959}$$

```{r}
(probability_ol * probability_ccl * probability_loan) / (probability_ol * probability_ccl * probability_loan + probability_onl * probability_ccnl * probability_nl)
```

Part F: The pivot table values from (B) is compared with Naive bayes probability.

From the Naive Bayes classifier, a higher value for P(Loan = 1 | CC
= 1, Online = 1) is obtained than with the direct computation obtained in part B. Interestingly, in part D we got the value of P(Loan = 1) as 0.096 same as value obtained in Naive Bayes classifier. So the concerned person being an online user with a bank-issued credit card is independent of the probability of the person accepting loan as suggested by Naive Bayes approach. 


Part G: Performing Naive Bayes on the data using the `naiveBayes()` function and assigning to the variable `naive_data`.
The value that is obtained for the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1) from  Naive bayes as seen is 0.09591693, which is equal to the value derived in part E.
```{r}
naive_data <- naiveBayes(`Personal Loan`~Online+CreditCard,data=Train_normalised)
naive_data
```


# AUC Value and ROC Curve
Plotting AUC (Area under curve) and ROC (Receiver Operating Characteristic) curve below using the `predict()` function by passing `naive_data` followed by using the `roc()` function and the `plot.roc()` (removing the second column).
```{r}
label_predicted <-predict(naive_data,Valid_normalised, type = "raw")
head(label_predicted)
```

```{r}
roc(Valid_normalised$Online, label_predicted[,2])
```
```{r}
plot.roc(Valid_normalised$Online,label_predicted[,2])
```