---
title: "Final project"
author: "Meenakshi Vaidhiyanathan"
date: "2024-04-21"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE )
library(readr)
library(class)
library(ISLR)
library(dplyr)
library(ggplot2)
library(caret)
library(pROC)
library(gmodels)
library(modeest)
library(moments)
library(ggcorrplot)
library(car)
library(DataExplorer)
library(skimr)
```

```{r}
Airlines <- read.csv("~/Desktop/Final project/Airlines.csv")
head(Airlines)
```

```{r}
dim(Airlines)
```

```{r}
str(Airlines)
```

missing values
```{r}
table(is.na(Airlines))
```

```{r}
colMeans(is.na(Airlines))
```


```{r}
Airlines<- Airlines[complete.cases(Airlines), ]
colMeans(is.na(Airlines))
```


```{r}
table(is.na(Airlines))
```

```{r}
NROW(Airlines)
```
Random sampling 
```{r}
set.seed(123)
x<- sample(1:103594, size = 10000, replace = FALSE)
```
New dataset after random sampling of 10000 rows
```{r}
Airlines<- Airlines[x,]
dim(Airlines)
```

```{r}
summary(Airlines)
```
```{r}
introduce(Airlines)
```
```{r}
skim(Airlines)
```


```{r}
mode_age<- mlv(Airlines$Age, method = "mfv")
mode_flightdist<- mlv(Airlines$Flight.Distance, method = "mfv")
print(paste("The mode of age is:",mode_age))
print(paste("The mode of flight distance in miles is:",mode_flightdist))
```

```{r}
quantile_Airlines_age<- quantile(Airlines$Age)
quantile_Airlines_age
```

```{r}
quantile_Airlines_flightDist<- quantile(Airlines$Flight.Distance)
quantile_Airlines_flightDist
```

```{r}
iqr_age<- IQR(Airlines$Age)
iqr_flightdist<- IQR(Airlines$Flight.Distance)
print(paste("The interqurtile range for age is:", iqr_age))
print(paste("The interqurtile range for flight distance is:", iqr_flightdist))
```

```{r}
sd_age<- sd(Airlines$Age)
sd_flightdist<- sd(Airlines$Flight.Distance)
print(paste("The standard deviation of age is:", sd_age))
print(paste("The standard deviation of flight distance is:", sd_flightdist))
```

```{r}
range_age<- range(Airlines$Age)
range_flightdist<- range(Airlines$Flight.Distance)
range_age
range_flightdist
```

```{r}
var_age<- var(Airlines$Age)
var_flightdist<- var(Airlines$Flight.Distance)
var_age
var_flightdist
```

```{r}
barplot(Airlines$Age, Airlines$Flight.Distance, col = "blue")
```


```{r}
hist(Airlines$Flight.Distance, col = "black", main = "Historgram of flight distance", breaks = 1000, xlab = "Flight distance")
```



```{r}
ggplot(data = Airlines, aes(x = Gender, y = Flight.Distance, color = "black")) +
  geom_point() +
  labs(title = "Flight distance vs. Gender",
       x = "Gender",
       y = "Flight distance")
```
Numerical vairables of dataset
```{r}
num_matrix <- Airlines[c("Age","Flight.Distance", "Departure.Delay.in.Minutes", "Arrival.Delay.in.Minutes")]
head(num_matrix)
```

Removing redundant variables
```{r}
Airlines <- Airlines[, !(names(Airlines) == "X")]
head(Airlines)
```

Removing redundant variables
```{r}
Airlines <- Airlines[, !(names(Airlines) == "id")]
head(Airlines)
```

Pairs for numerical variables 
```{r}
pairs(num_matrix)
```


```{r}
library(psych)
describe(Airlines)
```

```{r}
cor_matrix <- data.frame(cor(num_matrix))
head(cor_matrix)
```


```{r}
Airlines$Index <- 1:nrow(Airlines) 
head(Airlines)
```

```{r}
Airlines <- Airlines %>%
  select(ncol(Airlines), 1:(ncol(Airlines)-1))
head(Airlines)
```


```{r}
Gender_dummy<- ordered(Airlines$Gender)
Gender_dummy<- dummyVars(~Gender, data = Airlines)
gd<- predict(Gender_dummy, Airlines)
Class_dummy<- ordered(Airlines$Class)
Class_dummy<- dummyVars(~Class, data = Airlines)
cd<-predict(Class_dummy, Airlines)
Customer_dummy<- ordered(Airlines$Customer.Type)
Customer_dummy<- dummyVars(~Customer.Type, data = Airlines)
custd<-predict(Customer_dummy, Airlines)
Travel_dummy<- ordered(Airlines$Type.of.Travel)
Travel_dummy<- dummyVars(~Type.of.Travel, data = Airlines)
td<-predict(Travel_dummy, Airlines)
Satisfy_dummy<- ordered(Airlines$satisfaction)
Satisfy_dummy<- dummyVars(~satisfaction, data = Airlines)
sd<-predict(Satisfy_dummy, Airlines)
Airlines_dummy<- cbind(Airlines,gd,cd,custd,td,sd)
head(Airlines_dummy)
dim(Airlines_dummy)
```


```{r}
main_df<- data.frame(num_matrix, Airlines$Inflight.wifi.service, Airlines$Departure.Arrival.time.convenient, Airlines$Ease.of.Online.booking, Airlines$Gate.location, Airlines$Food.and.drink, Airlines$Online.boarding, Airlines$Seat.comfort, Airlines$Inflight.entertainment, Airlines$On.board.service, Airlines$Leg.room.service, Airlines$Baggage.handling, Airlines$Checkin.service, Airlines$Inflight.service, Airlines$Cleanliness, Airlines_dummy$GenderFemale, Airlines_dummy$ClassBusiness, Airlines_dummy$ClassEco, Airlines_dummy$`Customer.Typedisloyal Customer`, Airlines_dummy$`Type.of.TravelBusiness travel`, Airlines_dummy$satisfactionsatisfied)

colnames(main_df)<- c("Age", "Flight.Distance", "Departure.Delay.in.Minutes", "Arrival.Delay.in.Minutes", "Inflight.wifi.service", "Departure.Arrival.time.convenient", "Ease.of.Online.booking", "Gate.location", "Food.and.drink", "Online.boarding", "Seat.comfort", "Inflight.entertainment", "On.board.service", "Leg.room.service", "Baggage.handling", "Checkin.service", "Inflight.service", "Cleanliness", "GenderFemale", "ClassBusiness", "ClassEco", "Customer.Typedisloyal.Customer", "Type.of.TravelBusiness.travel", "satisfactionsatisfied")

head(main_df)
```

```{r}
dim(main_df)
```


```{r}
boxplot(main_df$Age, main= "Boxplot for age")
```

```{r}
boxplot(main_df$Flight.Distance, main= "Boxplot for flight distance")
```


Feature selection using ANOVA, PCA and ROC curve 
```{r}
lm_model <- lm(satisfactionsatisfied
 ~ ., data = main_df) 
anova_result <- anova(lm_model)
print(anova_result)
```


```{r}
colnames(anova_result)<- c("df", "Sum Sq", "Mean Sq", "f_value", "Pr_f")
print(anova_result)
``` 

```{r}
sorted_anova<- arrange(anova_result, desc(f_value))
print(sorted_anova)
```

pairs function 
```{r}
library(psych)
pairs.panels(main_df[,-24],
             gap=0,
             bg = c("red", "yellow", "blue")[main_df$satisfactionsatisfied],
             pch=21)
```

PCA method of feature selection 
```{r}
pc <- prcomp(main_df[,-24],
             center = TRUE,
            scale. = TRUE)
attributes(pc)
```

```{r}
print(pc)
```

```{r}
summary(pc)
```

visualisation of PCA
```{r}
# Perform PCA
pca_result <- prcomp(main_df, center = TRUE, scale. = TRUE)

# Load the factoextra package if not already loaded
library(factoextra)

# Visualize variables in the principal component space (biplot)
fviz_pca_var(pca_result, col.var = "cos2",
             gradient.cols = c("blue", "orange", "green"),
             repel = TRUE)
```


Corrplot
```{r}
cor_matrix <- data.frame(cor(main_df))
ggcorrplot(cor_matrix)
```


 

ROC scoring for feature selection
```{r}
#use roc_curve area as score
roc_imp <- filterVarImp(x = main_df[ ,1:24], y = main_df$satisfactionsatisfied)

#sort the score in decreasing order
roc_imp <- data.frame(cbind(variable = rownames(roc_imp), score = roc_imp[,1]))
roc_imp$score <- as.double(roc_imp$score)
print(roc_imp[order(roc_imp$score,decreasing = TRUE),])
```

```{r}
head(main_df)
```


After feature selection confirmation 
```{r}
main_df<- main_df[c("Online.boarding","Seat.comfort","On.board.service","Leg.room.service","Inflight.wifi.service","Type.of.TravelBusiness.travel","ClassBusiness","satisfactionsatisfied")]
```


```{r}
library(psych)
pairs.panels(main_df[,-8],
             gap=0,
             bg = c("red", "yellow", "blue")[main_df$satisfactionsatisfied],
             pch=21)
```

PCA confirmation 
```{r}
# Perform PCA
pca_result <- prcomp(main_df, center = TRUE, scale. = TRUE)

# Load the factoextra package if not already loaded
library(factoextra)

# Visualize variables in the principal component space (biplot)
fviz_pca_var(pca_result, col.var = "cos2",
             gradient.cols = c("blue", "orange", "green"),
             repel = TRUE)
``` 



##normalization
```{r}
norm_model<- preProcess(num_matrix, method=c("range")) 
model<- predict(norm_model, num_matrix)
```


```{r}
main_model<- data.frame(main_df) 

head(main_model)
```




```{r}
dim(main_df)
```


Corrlot after Feature selection
```{r}
corr_fe<- data.frame(cor(main_df))
ggcorrplot(corr_fe)
```


```{r}
## Partitioning the dataset
 
t<- createDataPartition(main_df$Online.boarding, p=0.6, list = FALSE)
train<- main_df[t, ]
test<- main_df[-t, ]
``` 


```{r}
train_predictors<- train[ ,1:7]
test_predictors<- test[ ,1:7]
```

##test label and train labels
```{r}
train_labels<-train[ ,8]
test_labels<-test[ ,8]
```


```{r}
search.grid<- expand.grid(k=c(2,7,9,15))
k_value<- train(satisfactionsatisfied ~ ., data = main_df, method= "knn", tuneGrid= search.grid)
k_value
```



```{r}
train_control <- trainControl(method = "cv", number = 100)
knn_model <- train(satisfactionsatisfied ~ ., data = main_df, method = "knn",
                   trControl = train_control,
                   tuneLength = 20)
print(knn_model)
```

##According to cross validation, k value =7 and implementation of knn
```{r}
predicted_test_lables<- knn(train_predictors, test_predictors, cl=train_labels, k=5)
head(predicted_test_lables)
```

##CrossTable(x=test_labels, y=predicted_test_lables, prop.chisq = FALSE)

```{r}
CrossTable(x=test$satisfactionsatisfied, y=predicted_test_lables, prop.chisq = FALSE)
```


##naive bayes
```{r}
library(e1071)
nb_model<- naiveBayes(satisfactionsatisfied~ ., data=train)
predicted_test_lables_nb<- predict(nb_model, test)
```


```{r}
CrossTable(x=test$satisfactionsatisfied, y=predicted_test_lables_nb, prop.chisq = FALSE)
```


```{r}
predicted_test_lables_nb<- predict(nb_model, test, type = "raw")
dim(predicted_test_lables_nb)
```



```{r}
roc(test$satisfactionsatisfied, predicted_test_lables_nb[,1])
```

```{r}
plot.roc(test$satisfactionsatisfied, predicted_test_lables_nb[,1])
```

