---
title: "Assignment 1"
author: "Meenakshi Vaidhiyanathan"
date: "2024-02-05"
output: pdf_document
---

## Dataset: 
The dataset used for this project can be found here:  
<https://www.kaggle.com/datasets/yasserh/housing-prices-dataset>.


## Setup: 

Libraries used :  

```{r setup, include=TRUE, echo=FALSE}
library(dplyr) 
library(ggplot2)
```

Loading the dataset : 
```{r load, include=TRUE}
housingData <- read.csv("./housing.csv")
head(housingData)
```
  
  
## Descriptive statistics for quantitative variables:
  
  
The `beds` variable includes results on the number of houses with their respective bedrooms. For eg: 136 houses have 2 bedrooms. 
```{r bedrooms}
beds<-table(housingData$bedrooms) 
head(beds)                      
```
  
  
The `bath` variable includes results on the number of houses with their respective bathrooms. For eg: 401 houses have 1 bathroom. 
```{r bathrooms}
bath<-table(housingData$bathrooms)
head(bath)
```
  
  
The `stories` variable includes results on the number of houses that are single, double, triple or quadruple storied building. For eg: 227 housese have are single storied. 
```{r stories}
stories<-table(housingData$stories)
head(stories)
```
  
  
The code below calculates the average price and the median value of the price of the houses.
```{r Price}
mean(housingData$price)
median(housingData$price)
```
  
  
The `order()` function is used on the dataset housingData to express values of area in descending order. The `head()` function is then used to print the valueof the house with highest area while the `tail()` function is used to print the value of the house with lowest area. 
```{r area}
highestAreaTable <-housingData[order(housingData$area,decreasing = TRUE),]
highAreaInfo<-head(highestAreaTable,1) 
lowAreaInfo<-tail(highestAreaTable,1)
highAreaInfo
lowAreaInfo
```
  
  
The `park` variable is used to store the number of houses with their respective parking facilities. For eg: 299 houses have zero parking facility. 
```{r parking}
park<-table(housingData$parking)
park 
```


### Descriptive statistics for categorical variables
  
  
The `num_ac` variable stores the number of houses with or without air conditioning. 
```{r airconditioning}
num_ac<-table(housingData$airconditioning)
num_ac
```
  
  
The `num_furnishing` variable stores the number of houses that are furnished, semi-furnished and unfurnished.  

```{r furnishing}
num_furnishing<-table(housingData$furnishingstatus)
num_furnishing
```
  
  
The `num_hotwaterheating` variable stores the number of houses that are with or without hot water heating facility. 
```{r hotwaterheating}
num_hotwaterheating<-table(housingData$hotwaterheating)
num_hotwaterheating
```

  
  
The `num_mainroad` variable stores the number of houses that are on the mainroad. 
```{r mainroad}
num_mainroad<-table(housingData$mainroad)
num_mainroad
```
  
  
The following code lists houses which have both air conditioning and hot water heating facilities. 
The `filter()` function filters the row of such houses that abide by the two conditions. The function was used as it supports conditions across multiple columns. 
```{r aircon and hotwater}
print(filter(housingData,(housingData$airconditioning== 'yes' & housingData$hotwaterheating=='yes')))
```
  
  

## Transform Variable
  
  
The dataset contained the absolute pricing of houses. It was transformed to be expressed as a multiple of 100,000. 
```{r transform, echo=TRUE}
priceperhunk<-housingData$price/100000
head(priceperhunk, 30)
```


## Plots

### Plotting quantitative variable:
  
  
The following histogram `hist()` plots the number of houses in the dataset vs the number of bedrooms.
```{r bedroomsplot, echo=FALSE}
# main() : the title of the graph
# col : color of the plot
# xlab : title of the x-axis
hist(housingData$bedrooms, main ="Number of houses with bedrooms", col= "purple", xlab = "bedrooms")
```
  
  
The following histogram `hist()` plots the number of houses in the dataset vs the number of parking spots.
```{r parkingspotplot, echo=FALSE}
# main() : the title of the graph
# col : color of the plot
# xlab : title of the x-axis
hist(housingData$parking, main ="Number of houses with parking", col = "yellow", xlab="parking")
```



### Plotting a scatter plot
  
  
The following plot depicts the price of the house in the dataset vs the number of bedrooms. The plot illustrates the variability in price based on the number of bedrooms.  

```{r bedroomsscatterplot, echo=FALSE}
# main() : the title of the graph
# col : color of the plot
# xlab : title of the x-axis
# ylab : title of the y-axis
plot(housingData$price/100000,housingData$bedrooms, main="Scatter plot of price vs bedrooms", col="blue", xlab="Price", ylab="Bedrooms")
```

  
  
The following plot illustrates the area of the house vs the number of parking spots available.  

```{r areaparkingplot, echo=FALSE}
# main() : the title of the graph
# col : color of the plot
# xlab : title of the x-axis
# ylab : title of the y-axis
plot(housingData$area,housingData$parking, main = "Scatter plot of Area vs Parking", col="orange", xlab = "Area", ylab = "Parking")
```


### Plotting of a categorical variable
  
  
- The following plot illustrates a categorical value i.e. `hotwaterheating` of values yes or no. 
- The function `ggplot()` was used as it allowed for plotting of categorical data. The `geom_bar()` function was used to depict a bar graph. The `aes()` or aesthetic function was needed to specify the categorical variable against the axis of the graph. Documentation for the `aes()` function can be found at <https://www.rdocumentation.org/packages/ggplot2/versions/3.4.4/topics/aes>. 

```{r hotwaterheatingplot, echo=FALSE}
ggplot(housingData, aes(x=hotwaterheating)) + geom_bar()
```
  
  
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
